The evolution of gene expression in the term placenta of viviparous mammals

### 2016-04-19 manuscript/manuscript.Rnw
fix grammar error in abstract
switch to SINE repeat from Alu repeat

  \parttitle{Conclusions} A diverse array of genes are differentially
  expressed and are likely to play a role in the anatomical
  variability of the placenta in the placentas of mammalian lineages.
  Furthermore, multiple genes which are associated with pre-eclampsia
  (\textit{CRH}, \textit{ADAM12}, \textit{KISS1}, \textit{PAPPA},
  \textit{IL1RL1}, and \textit{PLAC1}) show increased expression on
  the human lineage. Still, using the strictest filtering strategy the
  expression we find that a core set of only
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  genes is essential for the proper functioning of the placenta in all
  species analyzed.

@@

We found evidence of high expression in two novel expansions of genes
in the \elephant genome. One family includes 11 genes that are related
to the \textit{PRSS16} gene, a thymus-specific serine protease that
may be involved regulating the presentation of self-peptides in CD4+
T-lymphocytes \cite{Fornari.ea2012:TranscriptionprofilingofPrss16}.
All 11 of these genes are expressed in the elephant placenta, three of
which have very abundant mRNAs
(\cref{tab:top50_elephant_placenta} and \cref{fig:pag2_prss16_ctsl_psg_evolution_tree}C). It is
possible that this elephant-specific gene family may be involved in
immunity in the elephant placenta and the mediation of maternal-fetal
immune tolerance; however, future work will need to test these
hypotheses. Three other genes abundantly expressed in the elephant
placenta are homologues of \textit{SSLP1}, a gene with predominant
expression in the seminal vesicle in mouse
\cite{Li.ea2006:SSLP-1secretedLy-6protein}. Homologues of
\textit{SSLP1} are also expressed in \cow, \dog, \horse, and \sheep.
\textit{SSLP1} also contains an Alu-like retrotransposon and may be
involved in the development of villi via cell proliferation
\cite{Ushizawa.ea2009:Characterizationandexpressionanalysis}. This
gene is not found in primates. Expansion of these genes with high
expression in \elephant may indicate an elaboration of this system.

>>>>

  \parttitle{Conclusions} A diverse array of genes are differentially
  expressed and are likely to play a role in the anatomical
  variability of the placenta in the placentas of mammalian lineages.
  Furthermore, multiple genes which are associated with pre-eclampsia
  (\textit{CRH}, \textit{ADAM12}, \textit{KISS1}, \textit{PAPPA},
  \textit{IL1RL1}, and \textit{PLAC1}) show increased expression on
  the human lineage. Still, using the strictest filtering strategy
  we find a core set of only
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  genes is essential for the proper functioning of the placenta in all
  species analyzed.

@@

We found evidence of high expression in two novel expansions of genes
in the \elephant genome. One family includes 11 genes that are related
to the \textit{PRSS16} gene, a thymus-specific serine protease that
may be involved regulating the presentation of self-peptides in CD4+
T-lymphocytes \cite{Fornari.ea2012:TranscriptionprofilingofPrss16}.
All 11 of these genes are expressed in the elephant placenta, three of
which have very abundant mRNAs
(\cref{tab:top50_elephant_placenta} and \cref{fig:pag2_prss16_ctsl_psg_evolution_tree}C). It is
possible that this elephant-specific gene family may be involved in
immunity in the elephant placenta and the mediation of maternal-fetal
immune tolerance; however, future work will need to test these
hypotheses. Three other genes abundantly expressed in the elephant
placenta are homologues of \textit{SSLP1}, a gene with predominant
expression in the seminal vesicle in mouse
\cite{Li.ea2006:SSLP-1secretedLy-6protein}. Homologues of
\textit{SSLP1} are also expressed in \cow, \dog, \horse, and \sheep.
\textit{SSLP1} also contains an SINE repeat and may be
involved in the development of villi via cell proliferation
\cite{Ushizawa.ea2009:Characterizationandexpressionanalysis}. This
gene is not found in primates. Expansion of these genes with high
expression in \elephant may indicate an elaboration of this system.
